Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-03-02 manuscript/climate_sensitivity_draft.Rmd
#80

*Identification of drought years*

We identified droughts within the time period 1950-2009, defining drought [@slette_how_2019] as events with both anomalously dry peak growing season climatic conditions and widespread reductions in tree growth. Simultaneous consideration of both meteorological conditions and tree growth ensured that drought was the primary driver of observed growth declines and that our focus remained on droughts that substantially impacted the forest community.

We identified the years with driest conditions during May-August (MJJA), which stood out in the analysis of @helcoski_growing_2019 as the current-year months to which annual growth was most sensitive for trees at this site. We considered two metrics of moisture deficit: NOAA Divisional Data’s Palmer Drought Severity Index (PDSI) and the difference between monthly potential evapotranspiration ($PET$) and precipitation ($PRE$). These data were obtained from the ForestGEO Climate Data Portal (https://github.com/forestgeo/Climate) in August 2018, with monthly PET and PRE sourced from Climatic Research Unit high-resolution gridded dataset (CRU TS v.4.01; @harris_updated_2014). The driest years were identified through ranking mean MJJA $PET-PRE$ or $PDSI$ for the time period from driest to wettest. 
Three of the five years between 1950 and 2009 with greatest moisture deficit ($PET-PRE$) during MJJA consistently ranked as the three driest in terms of $PDSI$: 1966, 1977, and 1999, which had mean MJJA $PET-PRE$ of 83, 87, and 80 mm mo<sup>-1</sup>, respectively (Table S3). The years 1964 and 2007 also ranked among the five lowest $PET-PRE$ (84 and 82 mm mo<sup>-1</sup>, respectively), but were not among the lowest in terms of PDSI and were thus not identified as candidate years for inclusion as top drought years (Table S3).

>>>>

*Identifying drought years*

We identified droughts within the time period 1950-2009, defining drought [@slette_how_2019] as events with both anomalously dry peak growing season climatic conditions and widespread reductions in tree growth. Simultaneous consideration of both meteorological conditions and tree growth ensured that drought was the primary driver of the observed growth declines and that our focus remained on droughts that substantially impacted the forest community.

We identified the years with driest conditions during May-August (MJJA), which stood out in the analysis of @helcoski_growing_2019 as the months of the current year to which annual tree growth was most sensitive at this site. We considered two metrics of moisture deficit: NOAA Divisional Data’s Palmer Drought Severity Index (PDSI) and the difference between monthly potential evapotranspiration ($PET$) and precipitation ($PRE$). These data were obtained from the ForestGEO Climate Data Portal (https://github.com/forestgeo/Climate) in August 2018, with monthly PET and PRE sourced from Climatic Research Unit high-resolution gridded dataset (CRU TS v.4.01; @harris_updated_2014). The driest years were identified through by ranking mean MJJA PET-PRE or PDSI for the time period from driest to wettest. Three of the five years with greatest MJJA moisture deficit (PET-PRE) between 1950 and 2009 with greatest moisture deficit ($PET-PRE$) during MJJA consistently ranked as the three driest in terms of $PDSI$: 1966, 1977, and 1999, which had mean MJJA $PET-PRE$ of 83, 87, and 80 mm mo<sup>-1</sup>, respectively (Table S3). The years 1964 and 2007 also ranked among the top five in terms of moisture deficit lowest ($PET-PRE$ of 84 and 82 mm mo<sup>-1</sup>, respectively), but were not among the lowest in terms of PDSI and were thus not identified as candidate years for inclusion as top drought years (Table S3).
